Density-dependent selection in evolutionary genetics: a lottery model of Grime's triangle

### 2017-04-26 manuscript.tex
latest

The ideal environment is characterized by the near-absence of stress and disturbance: $d\ll 1$, whereas $b$ is potentially much larger than $1$. Disturbed environments are characterized by short bursts of high extrinsic mortality $d$ caused by physical destruction. We assume that in the disturbance extreme of Grime's triangle, many such bursts (which could each be be spatially localized) will occur each iteration (i.e. over the time from birth to reproductive maturity), and so we can approximate extreme disturbance by high constant mortality rates $d_i$, which are now extended to affect juveniles as well. We assume that disturbances are equally damaging to adults and juveniles, so that only $(1-d_i)\Delta_+ n_i$ rather than $\Delta_+ n_i$ territories are secured by genotype $i$ each iteration. Disturbed environments then correspond to $d_i$ being close to $1$ for all genotypes (most adults and juveniles are killed over an iteration).

@@

The alternative view is that stressed environments are highly competitive \citep{taylor_1990}. In particular, if consumable resources are scarce, we expect intense resource competition (for empirical support, see \citealt{davis_1998}). Stress does restrict growth and reproduction, but it also means that fewer individuals can be supported per unit area. Thus, stressed populations are actually at high density relative the environmental carrying capacity (but low density relative to ideal environments).

>>>>

The ideal environment is characterized by the near-absence of stress and disturbance: $d\ll 1$, whereas $b$ is potentially much larger than $1$.

Disturbed environments are characterized by short bursts of high extrinsic mortality $d$ caused by physical destruction. We assume that in the disturbance extreme of Grime's triangle, many such bursts (which could each be be spatially localized) will occur each iteration (i.e. over the time from birth to reproductive maturity), and so we can approximate extreme disturbance by high constant mortality rates $d_i$, which are now extended to affect juveniles as well. We assume that disturbances are equally damaging to adults and juveniles, so that only $(1-d_i)\Delta_+ n_i$ rather than $\Delta_+ n_i$ territories are secured by genotype $i$ each iteration. Heavily disturbed environments then correspond most adults and juveniles being killed every an iteration ($d_i\approx 1$).

@@

The alternative view is that stressed environments are highly competitive. In particular, if consumable resources are scarce, we expect intense resource competition (for empirical support, see \citealt{davis_1998}). Stress does restrict growth and reproduction, but it also means that fewer individuals can be supported per unit area. Thus, in this view, stressed populations are actually at high density relative the environmental carrying capacity (but low density relative to ideal environments; \citealt{taylor_1990}). Territorial contests are important.
